---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# easydb

<!-- badges: start -->

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)

<!-- badges: end -->

The goal of easydb is to simplify the process of connecting to databases using R

## Installation

You can install the development version of easydb like so:

``` r
# install.packages('remotes')
remotes::install_github('selkamand/easydb')
```

## Connecting to a database

### Functions

Connect to databases:

`easydb_connect('database_name')`

Disconnect from databases:

`easydb_disconnect(connection)`

### What to expect

The first time you try to connect to a database, you may have to answer some questions about the type of database (depending on what dbtype you're connecting to).

Once your configuration and credentials are set, you will not have to re-enter them unless you set the argument `from_scratch = TRUE` (which deletes your existing configuraion and credentials for the given database and prompts you to supply updated information.

## Quick start

```{r sqlite, eval=FALSE}
library(easydb)

# sqlite
sqlitedb <- system.file(package="easydb", "testdbs/mtcars.sqlite")
sqlite_connection <- easydb_connect(sqlitedb)

# mysql
# There is a public rfam mysql database we can use to test
# host: mysql-rfam-public.ebi.ac.uk
# username: rfamro
# port: 4497
# password: none
# database: Rfam
rfam_connection <- easydb_connect(dbname = "Rfam")


# postgresql
# RNAcentral hosts a public postgres database we can use for testing
# Hostname: hh-pgsql-public.ebi.ac.uk
# Port: 5432
# Database: pfmegrnargs
# User: reader
# Password: NWDMCE5xdipIjRrp
rna_central_connection <- easydb_connect(dbname = 'pfmegrnargs')


# Don't forget to disconnect from databases when you finish!
easydb_disconnect(sqlite_connection)
easydb_disconnect(rna_central_connection)
easydb_disconnect(rfam_connection)
```

## Disconnect

Once you've finished working with a database, its best to disconnect from the db using `easydb_disconnect(connection_object)`

# How does easydb work?

Database configurations (host, port, driver, etc) are stored in a configuration file that will be created at \~/.easydb or some custom user-specified location.

Credentials are stored separately in your operating systems credential store. This is powered by `keyring`. If you're on linux you may need to install the secret service library. See the [readme](https://github.com/r-lib/keyring) for details
